---
title: "Principle Component Aanalysis - Fernando Carnauba fa2529"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
rm(list = ls())
D1 <- read.csv("~/R/HUDK/assignment4/Assistments-confidence.csv")

#We won't need the id variable, so remove that.
D1 <- dplyr:: select(D1, -id) 

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later

# ok, let's save it:
jpeg('corr1.jpg')
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
dev.off()

```

## Create a new data frame with the mean_correct variables removed

```{r}
D2 <- dplyr:: select(D1, -mean_correct) 

#Then, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)

```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Decide which components you would drop and remove them from your data set.

ANSWER: The whole point of PCA is to reduce the amount of data we are dealing with. So, my first idea is to really "test" the PCA model by reducing the number dimension by at least half. Also, looking into components variances, we can see the second and thrid components have quite similar predictive power, with a significant drop after that (the steepest drop in explanatory power is exctually from PC1 to PC2, but running the whole analysis with one PC would be simpler and might take away the pegadogical intention of this exercise - also, PC1 alone only accounts for 35% of total variance). So I will keep componets 1-3, with a total "explained variance" of about 63%. 

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x[,1:3])

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)
COR2

corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

```

Well, "mean_correc"t correlates positively with PCs 1 and 2, but not with PC3. But I'm not sure how to find out whether we lost important information about mean_data only from this analysis. I will try to answer this with an addition analysis (below). 

## Did we loose predictive power (with regards to mean_correct)?

To check if we lost predictive information about "mean_correct" that was contained in the other initial 6 variables, in comparison to the 3 components, I will look into the R2 of linear regressions (my background is in Economics - so I recognize I may be somewhat addicted to regressions..!). If we did not loose a lot of information, then the R2 of the linear regression of "mean_correct" on the 6 variables shouldn't be much higher than that of the linear regression of "mean_correct" on the 3 componets! 

```{r}

# Regression of mean_correct on the 6 variables (quick note: we don't even need to rescale, because that doesn't change the R2 in regression analysis...)
  
model1 <- lm(mean_correct ~ prior_prob_count + prior_percent_correct + problems_attempted + mean_hint + mean_attempt + mean_confidence, data = D1)
  
# Regression of mean_correct on the 3 componets
model2 <- lm(D1$mean_correct ~ PC1 + PC2 + PC3, data = D4)

summary(model1)
summary(model2)

```

well, our R2 did become quite smaller, from .39 in the first model to .22 in the second. We might have lost relevant information with the dimensionality reduction!

## Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

#Now examine your components and try to come up with substantive descriptions of what some might represent?


#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

```
## Interpreting PCs in Part 2

## PC1 - Students struggling
PC1 seems to mainly take into account information related to struggling in answering the questions, since it has high values for the variables "mean_attemp" and "mean_hint" . Also, they all have the same sign (negative), which indicates that they are all taken into account in the same direcetion by this component. The higher the mean of attempts needed to solve a problem in this session (mean_attempt), the higher the change this person is struggling. Also, the higher the frenquecy of asking for hints (mean_hint), the higher the probability of struggling. 

## PC2 - Sudent prior achievement
This component seems to mostly capture students prior knowledge/achievement, with high values for "prior_prob_count" and "prior_percent_correct". The higher the number of problems a student has already anwered in the system (prio_prob_count), the higher the likelihood of having strong previous knowledge. For the case of percentage of prior correct answers (prio_prob_count) the relationship is even more direct - a higher percentage of previous correct answers strongly signals that the student was doing well in the previous sections.

## PC3 - Student confidence
This component seems to rely much on student confidence. It makes sense that this variable stands out "individually" in one of the eigenvectors, since it is almost uncorrelated with any other variable (notice in the correlation matrix that its maximum correlation with any other variable is .08).




# Part III  

Also in this repository is a data set collected from class that shows how many students thought that a TC program was related to andother TC program. Use PCA to look for components of related programs. Explain why you think there are relationships between these programs.

```{r}

# Again, let's first look into the correlations between variables (i.e., which programs go "together" as related/unrelated to other programs)

TCprograms <- read.csv("~/R/HUDK/assignment4/TCprograms.csv", sep="")
TC1 <- dplyr:: select(TCprograms, -program) #get rid of program name
TC1 <- dplyr:: select(TC1, -cluster)[1:68,]  #get rid of cluster (not suitable fot PCA, which is a cardinal analysis)

TC2 <- scale(TC1, center = TRUE) #rescale

# Generate pairwise correlations
TCcor <- cor(TC2)

# This is quite a large matrix. I tried to visually inspect it and it was pretty hard to find any obvious relationships (like one program that correlates strongly with many others). What is reassuring, though, is that there exist a few "large" values (above .3, for example) that stand out from a relative overall low correlation level between of programs. Those "high values" mean that the two programs usually "go in tandem" as either being rated as "similar" or "not similar" to each of the other remaining programs. We will keep this data matrix of correlations saved, as it might come to help later.

# I decided to also try something "simpler" as an initial exploratory analysis. For each program, will find what was rated directly by students as its most similar program. 

closest <- data.frame(TCprograms$program[1:68],colnames(TC2)[max.col(TC2)],apply(TC1,1,max))
colnames(closest) <- c("program","closest","n_paired")

#now we can see which programs have a "very" similar pair (I picked the rule "student responses >= 8" after visually inpecting the data)

closest[closest$n_paired >= 8,]

```

Here are my from "gotchas" this initial analysis:

1. A "Technology and Education" cluster. It seems that "Communication.Media.and.Learning.Technologies", "Design.and.Development.of.Digital.Games" and "	Creative.Technologies" are related (makes sense to me...)
2. A Nutrition pair. "Diabetes.Education" and "nutrition" (definitely makes sense..)
3. A Math/Stat pair. "Mathematics" and "Measurement..Evaluation.and.Statistics" (not true, in my opinion...)
4. A Reading pair. "Reading" and "Literacy" are also viewed as similar (makes sense to me...)


Ok, time for some PCA analysis

```{r}
# run PCA
TCpca <- prcomp(TC2, scale = TRUE)

#let's see how each component accounts for the variance. I'm not going to print the summary here because it is really long. But I had I did have a look into it and I saw that only the first 4 components contribute to explaining more than 5% of the variance each. 

# So let's print the eigenvector for those first 4 PCs and let's try to interpret them. 

TCpca$rotation[,1:4]

```


## PC1 - Social Sciences and Leadership dimension

We print below the 10 programs with largest positive weight for PC1 (honestly, looking into negative weights seems quite weired here - there is no meaning in being "oposite" to a given subject, and I decided to restrict my analysis to postive coeficients).

This PC gives a lot of weight to social sciences and leadership-related (including school administration) programs. Every single subject among the 10 with largest positive weight falls into such areas. Interestigly, this dimension has nothing to do with our initial exploratory analysis!(But hold on, Component 3 does have something to do with it!)


```{r}

names(TCpca$rotation[order(-TCpca$rotation[,1]),1][1:10])

```

## PC2 - Psychology-and-Health dimension

We print below the 10 programs with largest positive weight for PC2.

This gives a lot of weight to programs related to health and psychology. Every very single subject among the 10 with largest positive weight falls into such category - except for the 10th largest, "Physical Education" (this is not too fal from the "Heath" topic, though - is it?). 

```{r}
# 10 programs with largest positive weight for PC2.

names(TCpca$rotation[order(-TCpca$rotation[,2]),2][1:10])

```

## PC3 - The-future-of-Ed-dimension (hooray!)
We print below the 10 programs with largest positive weight for PC3.

We finally found something that has some relationship with our initial exploratory analysis. This dimension includes in its 10 most relevant positive variables almost exclusively technology-related programs. The only exceptions here are "cognitive science" and "mathematics". 

```{r}
# 10 programs with largest positive weight for PC3.

names(TCpca$rotation[order(-TCpca$rotation[,3]),3][1:10])

```

## PC4 - The non-teacher-education-dimension
 
This componet seemed quite hard to interpretate, as it shows a combination of elements related to the dimensions PC1 and PC2 such as  "Leadership", "Sociology", "Politics", "Change.Leadership", "Physiology", "Economics.and.Education", "Health.Education", and"Nursing". In this case, I did get curious to see what were the largest negative weights. Interestingly, some coherence around the topic of "tecaher education" shows up. This is the case for "English.Education", "Teaching.English", "Literacy", "Bilingual.Bicultural.Education", "Reading", and "Early.Childhood.Special.Education". So I suggest that this dimension might be interpreted as the "non-teacher-education" dimension! To me, this actually makes some sense once we remember that this data is from Teachers College. At this school Teacher Ed. vs. Non-Techer Ed. is indeed a meaningful divider.

We print below the 10 programs with largest positive weight for PC4. We also do the same for the 10 larges negative weights.  

```{r}
# 10 programs with largest positive weight for PC4.
names(TCpca$rotation[order(-TCpca$rotation[,4]),4][1:10])

#10 programs with largest NEGATIVE weight for PC4.
names(TCpca$rotation[order(TCpca$rotation[,4]),4][1:10])

```



