---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv")
library(tidyverse)
  #We won't need the id variable, so remove that.
D1<-D1[-1]

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)
install.packages("corrplot")
library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
```

## Create a new data frame with the mean_correct variables removed

```{r}
D2 <- D1[-4]

#Then, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Think about which components you would drop and make a decision
Looking at the Scree plot above, the first two components account for a relatively large amount of variance compared to the last four components. However, the third component also has a variance value of over 1 so that should mean it explains slightly more variance than just a single item in our dataset. So I would probaby keep 3 components.

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?
```

#No, it doesn't look like I would have lost very much important information. The 2nd through 6th components all plateaued at a variance of 1. They would not have contributed much to explaining the variance in our data.

```{r}
COR2 <- cor(D4)
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Then, scale and center your data for easier interpretation
D5 <- scale(D4, center = TRUE)

pca <- prcomp(D5, scale = TRUE)

plot(pca, type = "lines")

```
## Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
eigen(COR2)

```


```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

load1<-sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

#Now examine your components and try to come up with substantive descriptions of what some might represent?
```
The biplot was sort of helpful in understanding which components would be helpful. What is confusing to me is that PC1 and PC2 are supposed to be the x- and y-axes but there is also a vector representation of them (not pointing directly along the axes). Anyway, what I think it shows is that PC1 "goes together" well with the mean_correct variable. I'm unsure why the actual arrow of PC1 points in the negative direction of its own axis.

PC2 and PC4 are the next closest components to pointing in the same direction as mean_correct. So perhaps it would be best to keep PC1,PC2, and PC4. Although, PC3 and PC5 are aligned with the y-axis (PC2) so they are highly correlated to PC2.

From the initial correlation plot, "prior_percent_corret" had a high correlation with "mean_correct" so it is probably represented in PC1 or PC2. "Problems_attempted" and "prior_prob_count" had similarly small correlations with "mean_correct" so they must be represented by PC1 and PC2 as well. Since "mean_hints" has such a negative correlation with "mean_correct" I would assume that it is mostly represented by PC 6.

```{r}


#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

#I'm not sure if what I have below is what you want.

#PC1,2, & 4 are my composite variables that have values for each student
#remove PC3, PC5, and PC6 from the pca
D6<-select(D4,"PC1","PC2","PC4","D1$mean_correct")

COR3 <- cor(D6)
corrplot(COR3, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

```
# Part III  
## Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

```{r}
E1<-read.csv("humor_data.csv")

#Pairwise correlations
E2<-E1[,-c(33:39)]
##Looking at which Qns best correlate with only the self-deprecating humor
#E2<-E1[,-c(33:35,37:39)] 

##Looking at which Qns best correlate with the four different kinds of humor
#E2<-E1[,-c(37:39)]
CORE <- cor(E2)

corrplot(CORE, order="AOE", method="circle", tl.pos="lt", type="upper",   tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#E3<-E2[,-33]

E3 <- scale(E2, center = TRUE)
pca <- prcomp(E3, scale = TRUE)
summary(pca)
plot(pca, type = "lines")

#So looking at the Scree plot and the summary, components 1-5 look good because they can explain variances from  more than 19% at best to a little above 4.5% at worst. After the fifth component the graph levels off and each component only accounts for 3% or less.

E4 <- as.data.frame(pca$x)

#Attach one of the variables for humor categories from the original data frame to E4.

#E5 <- cbind(E4, as.data.frame(E1$affiliative))  #AFFILIATIVE HUMOR
#E5 <- cbind(E4, as.data.frame(E1$selfenhancing))  #Self-enhancing HUMOR
#E5 <- cbind(E4, as.data.frame(E1$agressive))  #AGGRESSIVE HUMOR
E5 <- cbind(E4, as.data.frame(E1$selfdefeating))  #SELF-DEFEATING HUMOR


CORE2 <- cor(E5)
corrplot(CORE2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Then, scale and center your data for easier interpretation
E6 <- scale(E5, center = TRUE)

pca <- prcomp(E6, scale = TRUE)

plot(pca, type = "lines")

#get the eigenvectors
eigen(CORE2)

pca$rotation

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

load1<-sweep(loadings, 2, colSums(loadings), "/")

biplot(pca)

#The biplot excluding any of the humor types shows 4 groupings of questions going in opposite directions from each other.  I wonder if each of these clusters of Questions corresponds to one of the 4 categories of humor they listed?
#Doing PCA for just Affiliative humor shows we have one composite variable that can explain 54% of the variation.
#For Self-Enhancing humor, the PC1 composite variable represents 49%  of the variance.
#For Aggressive humor, the PC1 composite variable represents 49%  of the variance.
#For Self-deafeating humor, the PC1 composite variable represents 40% of the variance.



```

