---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

```{r setup}
library(dplyr)
library(plyr)

#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)
library(corrplot)
library(tidyr)
```

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv")
#We won't need the id variable, so remove that.
D1 <- D1 %>%
  select(-id)

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
```

## Create a new data frame with the mean_correct variables removed

```{r}
D2 <- D1 %>%
  select(-mean_correct)

#The, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)

```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
pca
pca$rotation
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Decide which components you would drop and remove them from your data set.
I would drop PC6, as it accounts for only ~9% of the variance.

```{r}
pca_df <- as.data.frame(pca$x)
D3 <- pca_df %>%
  select(-PC6)
```

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3

#Attach the variable "mean_correct" from your original data frame to D3.

D1$mean_correct <- scale(D1$mean_correct)

D4 <- cbind(D3, as.data.frame(D1$mean_correct))
            
#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)

corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

```
PC1, PC2 and PC4 have significant correlation with mean_correct. Hence removing those PCs may have lost important information on the mean_correct.


## Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

```

##Now examine your components and try to come up with substantive descriptions of what some might represent?

A large proportion of PC1 is mainly comprises  mean_hint, mean_attempt and problems_attempted, therefore correlated variables relating to the current session that the student is in. PC2 is largely made up of correlated variables relating to prior session, which are prior_prob_count and prior_percent_correct.

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  
biplot(pca)

## Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

```{r}

df_with_mean_correct <- cbind(pca_df, as.data.frame(D1$mean_correct))

df_with_mean_correct
#Note: mean_correct is column V1

COR2 <- cor(df_with_mean_correct)

corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

```
# Part III  

## Also in this repository is a data set collected from class that shows how many students thought that a TC program was related to andother TC program. Use PCA to look for components of related programs. Explain why you think there are relationships between these programs.

```{r upload and clean up TC data}

TC_data <- read.csv("TCprograms.csv", sep="")

rownames(TC_data) <- TC_data$pro

TC_data <- TC_data %>%
  select(-program)

View(TC_data)

```

```{r}
COR_TC <- cor(TC_data)

TC_data <- scale(TC_data, center = TRUE)

pca_TC <- prcomp(TC_data, scale = TRUE)

pca_TC$rotation
```
