---
title: "Principle Component Aanalysis"
output: html_document
---
#Data
The data you will be using comes from teh Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

#Start by uploading the data
```{r}
D1 <- read.table("~/Desktop/FALL_2016/Assignment4/Assistments-confidence.csv", sep = ",", header = TRUE)


  #We won't need to id variable, so remove that.

D2 <- dplyr::select(D1, 2:8)

```

#Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
```

#Create a new data frame with the mean_correct variables removed

```{r}
D2 <- dplyr::select(D1, -4)

#The, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

#Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
```

#Although the algorithm does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

[1] 1.2825140 1.0543565 1.0245688 0.9621486 0.8556715 0.7320146

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

[1] 1.6448423 1.1116675 1.0497412 0.9257299 0.7321737 0.5358454

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

Importance of components:
                          PC1    PC2    PC3    PC4    PC5     PC6
Standard deviation     1.2825 1.0544 1.0246 0.9621 0.8557 0.73201
Proportion of Variance 0.2741 0.1853 0.1750 0.1543 0.1220 0.08931
Cumulative Proportion  0.2741 0.4594 0.6344 0.7887 0.9107 1.00000

#We can lot this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

#Think about which components you would drop and make a decision

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")
D4 <- scale(D4, center = TRUE)
> pca2 <- prcomp(D4, scale = TRUE)
> pca2$sdev
[1] 1.2737535 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 0.6144527
> pca2$sdev^2
[1] 1.6224479 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 0.3775521
> summary(pca2)
Importance of components:
                          PC1    PC2    PC3    PC4    PC5    PC6     PC7
Standard deviation     1.2738 1.0000 1.0000 1.0000 1.0000 1.0000 0.61445
Proportion of Variance 0.2318 0.1429 0.1429 0.1429 0.1429 0.1429 0.05394
Cumulative Proportion  0.2318 0.3746 0.5175 0.6603 0.8032 0.9461 1.00000
> plot(pca2, type = "lines")  
```

```{r}
#so are we okay to remove prior-prob-count? 
D5 <- dplyr::select(D1, 1:7)
View(D5)
D5 <- dplyr::select(D1, 2:7)
COR3 <- cor(D5)
corrplot(COR3, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45,addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")
D5 <- scale(D5, center = TRUE)
COR3 <- cor(D5)
pca3 <- prcomp(D5, scale = TRUE)
pca3$sdev
[1] 1.3620848 1.1318578 1.0118542 0.8715683 0.8555761 0.5900275
pca3$sdev^2
[1] 1.8552750 1.2811020 1.0238488 0.7596313 0.7320104 0.3481324
summary(pca3)
Importance of components:
                          PC1    PC2    PC3    PC4    PC5     PC6
Standard deviation     1.3621 1.1319 1.0119 0.8716 0.8556 0.59003
Proportion of Variance 0.3092 0.2135 0.1706 0.1266 0.1220 0.05802
Cumulative Proportion  0.3092 0.5227 0.6934 0.8200 0.9420 1.00000
plot(pca3, type = "lines")
D6 <- as.data.frame(pca3$x)
View(D6)
D6 <- cbind(D5, as.data.frame(D1$prior_prob_count))
COR4 <- cor(D6)
Based on the graph, I think the variance gets worse so I should not remove the prior_prob_count
```

