---
title: "Principle Component Aanalysis"
output: html_document
---
#Data
The data you will be using comes from teh Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

#Start by uploading the data
```{r}
D1 <- read.table("~/Desktop/FALL_2016/Assignment4/Assistments-confidence.csv", sep = ",", header = TRUE)


  #We won't need to id variable, so remove that.

D2 <- dplyr::select(D1, 2:8)

```

#Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
```

#Create a new data frame with the mean_correct variables removed

```{r}
D2 <- dplyr::select(D1, -4)

#The, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

#Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
```

#Although the algorithm does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

[1] 1.2825140 1.0543565 1.0245688 0.9621486 0.8556715 0.7320146

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

[1] 1.6448423 1.1116675 1.0497412 0.9257299 0.7321737 0.5358454

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

Importance of components:
                          PC1    PC2    PC3    PC4    PC5     PC6
Standard deviation     1.2825 1.0544 1.0246 0.9621 0.8557 0.73201
Proportion of Variance 0.2741 0.1853 0.1750 0.1543 0.1220 0.08931
Cumulative Proportion  0.2741 0.4594 0.6344 0.7887 0.9107 1.00000

#We can lot this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

#Think about which components you would drop and make a decision

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")
D4 <- scale(D4, center = TRUE)
> pca2 <- prcomp(D4, scale = TRUE)
> pca2$sdev
[1] 1.2737535 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 0.6144527
> pca2$sdev^2
[1] 1.6224479 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 0.3775521
> summary(pca2)
Importance of components:
                          PC1    PC2    PC3    PC4    PC5    PC6     PC7
Standard deviation     1.2738 1.0000 1.0000 1.0000 1.0000 1.0000 0.61445
Proportion of Variance 0.2318 0.1429 0.1429 0.1429 0.1429 0.1429 0.05394
Cumulative Proportion  0.2318 0.3746 0.5175 0.6603 0.8032 0.9461 1.00000

> plot(pca2, type = "lines")  
```

```{r}
#so are we okay to remove prior-prob-count? 
D5 <- dplyr::select(D1, 1:7)
View(D5)
D5 <- dplyr::select(D1, 2:7)
COR3 <- cor(D5)
corrplot(COR3, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45,addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")
D5 <- scale(D5, center = TRUE)
COR3 <- cor(D5)
pca3 <- prcomp(D5, scale = TRUE)
pca3$sdev
[1] 1.3620848 1.1318578 1.0118542 0.8715683 0.8555761 0.5900275
pca3$sdev^2
[1] 1.8552750 1.2811020 1.0238488 0.7596313 0.7320104 0.3481324
summary(pca3)
Importance of components:
                          PC1    PC2    PC3    PC4    PC5     PC6
Standard deviation     1.3621 1.1319 1.0119 0.8716 0.8556 0.59003
Proportion of Variance 0.3092 0.2135 0.1706 0.1266 0.1220 0.05802
Cumulative Proportion  0.3092 0.5227 0.6934 0.8200 0.9420 1.00000
plot(pca3, type = "lines")
D6 <- as.data.frame(pca3$x)
View(D6)
D6 <- cbind(D5, as.data.frame(D1$prior_prob_count))
COR4 <- cor(D6)
Based on the graph, I think the variance gets worse so I should not remove the prior_prob_count
```

```{r}
Now print out the eigenvectors (often called loadings) for the components you generated:

pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

#Now examine your components and try to come up with substantive descriptions of what some might represent?

mean hint and mean attempt showes the most variance in PC1

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)
```

```{r}
#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

so you have to calculate with one student's value with
be more detail in terms of naming (reasoning behind it)
pc1=persistence 
pc2=priorknowledge 
pc3=confidence 
pc4=preattempt 
pc5=postattempt 
pc6=cheat
D4$persistence <- (0.12423113*D4$prior_prob_count) + (0.08035956*D4$prior_percent_correct) + (0.21744737*D4$problems_attempted)  + (0.30223780*D4$mean_hint) + (0.25863458*D4$mean_attempt) + (0.01708956*D4$mean_confidence)
D4$priorknowledge <- (0.25081186*D4$prior_prob_count) + (0.44677621*D4$prior_percent_correct) + (0.17344469*D4$problems_attempted)  + (0.06843387*D4$mean_hint) + (0.04658844*D4$mean_attempt) + (0.01394492*D4$mean_confidence)
D4$confidence <- (0.22101700*D4$prior_prob_count) + (0.05108998*D4$prior_percent_correct) + (0.20060288*D4$problems_attempted)  + (0.04415217*D4$mean_hint) + (0.02527878*D4$mean_attempt) + (0.45785919*D4$mean_confidence)
D4$preattempt <- (0.31516257*D4$prior_prob_count) + (0.12062699*D4$prior_percent_correct) + (0.14475664*D4$problems_attempted)  + (0.05129246*D4$mean_hint) + (0.14203987*D4$mean_attempt) + (0.22612148*D4$mean_confidence)
D4$postattempt <- (0.003664468*D4$prior_prob_count) + (0.153315014*D4$prior_percent_correct) + (0.303884750*D4$problems_attempted)  + (0.052483764*D4$mean_hint) + (0.357699023*D4$mean_attempt) + (0.128952980*D4$mean_confidence)
D4$cheat <- (0.14011651*D4$prior_prob_count) + (0.17770154*D4$prior_percent_correct) + (0.15748983*D4$problems_attempted)  + (0.35608836*D4$mean_hint) + (0.16165478*D4$mean_attempt) + (0.00694897*D4$mean_confidence)

                             PC1        PC2        PC3        PC4         PC5        PC6
prior_prob_count      0.12423113 0.25081186 0.22101700 0.31516257 0.003664468 0.14011651
prior_percent_correct 0.08035956 0.44677621 0.05108998 0.12062699 0.153315014 0.17770154
problems_attempted    0.21744737 0.17344469 0.20060288 0.14475664 0.303884750 0.15748983
mean_hint             0.30223780 0.06843387 0.04415217 0.05129246 0.052483764 0.35608836
mean_attempt          0.25863458 0.04658844 0.02527878 0.14203987 0.357699023 0.16165478
mean_confidence       0.01708956 0.01394492 0.45785919 0.22612148 0.128952980 0.00694897

#combined meancorrect variable from original data
D6 <- cbind(D4, as.data.frame(D1$mean_correct))
View(D6)

#created correlogram
COR4 <- cor(D6)
corrplot(COR4, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")

 it seems like there are high connection between cheat and post attempt, persistent and post attempt, and cheat and persistance. it seems like students got hint to solve the problem, being persistent on solving the problem using hint.
```

Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

#Start by uploading the data
```{r}
D7 <- read.table("~/Desktop/FALL_2016/Assignment4/humor_data.csv", sep = ",", header = TRUE)
D8 <- dplyr::select(D7,1:32)
```

# wanted to code the each questions name into my choice, but ended up using what professionals have found in their article
```{r}
Q1. I usually don?t laugh or joke around much with other people. =nolaugh #GN
Q2. If I am feeling depressed, I can usually cheer myself up with humor. =cheeruphumor #SE
Q3. If someone makes a mistake, I will often tease them about it. =teasemistake #AG
Q4. I let people laugh at me or make fun at my expense more than I should. =makefunmorethan #SD
Q5. I don?t have to work very hard at making other people laugh?I seem to be a naturally humorous person. =bornhumor #AF/SE
Q6. Even when I?m by myself, I?m often amused by the absurdities of life. = absurdities #SE
Q7. People are never offended or hurt by my sense of humor. =neveroffend #AF
Q8. I will often get carried away in putting myself down if it makes my family or friends laugh.=putmyselfdown #SD
Q9. I rarely make other people laugh by telling funny stories about myself. =neverfunny
Q10. If I am feeling upset or unhappy I usually try to think of something funny about the situation to make myself feel better. =putmyselfup
Q11. When telling jokes or saying funny things, I am usually not very concerned about how other people are taking it. =givenod
Q12. I often try to make people like or accept me more by saying something funny about my own weaknesses, blunders, or faults. =playingweak
Q13. I laugh and joke a lot with my closest friends. =insiderjoke
Q14. My humorous outlook on life keeps me from getting overly upset or depressed about things. =alwayspositive
Q15. I do not like it when people use humor as a way of criticizing or putting someone down. =nofun
Q16. I don?t often say funny things to put myself down. =dpromp
Q17. I usually don?t like to tell jokes or amuse people. =serious
Q18. If I?m by myself and I?m feeling unhappy, I make an effort to think of something funny to cheer myself up. =cheerupfun
Q19. Sometimes I think of something that is so funny that I can?t stop myself from saying it, even if it is not appropriate for the situation. =unapologetic
Q20. I often go overboard in putting myself down when I am making jokes or trying to be funny. =overboard
Q21. I enjoy making people laugh. =makelaugh
Q22. If I am feeling sad or upset, I usually lose my sense of humor. =losehumor
Q23. I never participate in laughing at others even if all my friends are doing it. =neverlaugh
Q24. When I am with friends or family, I often seem to be the one that other people make fun of or joke about.=madefunof
Q25. I don?t often joke around with my friends.=nojokewf
Q26. It is my experience that thinking about some amusing aspect of a situation is often a very effective way of coping with problems.=amusingcope
Q27. If I don?t like someone, I often use humor or teasing to put them down.=sneakyjoke
Q28. If I am having problems or feeling unhappy, I often cover it up by joking around, so that even my closest friends don?t know how I really feel. =maskjoke
Q29. I usually can?t think of witty things to say when I?m with other people. =nowitty
Q30. I don?t need to be with other people to feel amused ? I can usually find things to laugh about even when I?m by myself. =alonefun #SE
Q31. Even if something is really funny to me, I will not laugh or joke about it if someone will be offended. =cautious #AF
Q32. Letting others laugh at me is my way of keeping my friends and family in good spirits. =goodspirit #AF
names(D8)<-c("Q1_AF", "Q2_SE", "Q3_AG", "Q4_SD", "Q5_AF", "Q6_SE", "Q7_AG", "Q8_SD", "Q9_AF", "Q10_SE","Q11_AG","Q12_SD","Q13_AF","Q14_SE","Q15_AG","Q16_SD","Q17_AF","Q18_SE","Q19_AG","Q20_SD","Q21_AF","Q22_SE","Q23_AG","Q24_SD","Q25_AF","Q26_SE","Q27_AG","Q28_SD","Q29_AF","Q30_SE","Q31_AG","Q32_SD")
```

#Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
library(corrplot)

#Generate pairwise correlations
COR8 <- cor(D8)

corrplot(COR8, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
There were not strong correlationship between each questions. But still, between questions which were grouped into same category, they showed medium correlation.
```

#Now run the PCA on the new data frame

```{r}
pca8 <- prcomp(D8, scale = TRUE)
```

#Although the algorithm does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca8$sdev

 [1] 2.4943111 1.8024083 1.6831741 1.4784656 1.2030427 1.0125176 1.0014851 0.9583264 0.9193296 0.8934064 0.8821141 0.8531886
[13] 0.8449333 0.8229371 0.7934106 0.7797708 0.7663446 0.7642473 0.7465136 0.7287594 0.7236779 0.7120325 0.7065632 0.6914301
[25] 0.6781329 0.6578595 0.6531532 0.6458890 0.6137997 0.6023745 0.5625567 0.5278429

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca8$sdev^2
[1] 6.2215881 3.2486758 2.8330752 2.1858605 1.4473118 1.0251919 1.0029725 0.9183895 0.8451669 0.7981749 0.7781253 0.7279308
[13] 0.7139123 0.6772255 0.6295003 0.6080425 0.5872840 0.5840739 0.5572825 0.5310902 0.5237097 0.5069903 0.4992316 0.4780756
[25] 0.4598642 0.4327792 0.4266091 0.4171726 0.3767501 0.3628550 0.3164700 0.2786182

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca8)

Importance of components:
                          PC1    PC2     PC3     PC4     PC5     PC6     PC7    PC8     PC9    PC10    PC11    PC12    PC13    PC14
Standard deviation     2.4943 1.8024 1.68317 1.47847 1.20304 1.01252 1.00149 0.9583 0.91933 0.89341 0.88211 0.85319 0.84493 0.82294
Proportion of Variance 0.1944 0.1015 0.08853 0.06831 0.04523 0.03204 0.03134 0.0287 0.02641 0.02494 0.02432 0.02275 0.02231 0.02116
Cumulative Proportion  0.1944 0.2959 0.38448 0.45279 0.49802 0.53005 0.56140 0.5901 0.61651 0.64145 0.66577 0.68851 0.71082 0.73199
                          PC15   PC16    PC17    PC18    PC19   PC20    PC21    PC22   PC23    PC24    PC25    PC26    PC27    PC28
Standard deviation     0.79341 0.7798 0.76634 0.76425 0.74651 0.7288 0.72368 0.71203 0.7066 0.69143 0.67813 0.65786 0.65315 0.64589
Proportion of Variance 0.01967 0.0190 0.01835 0.01825 0.01742 0.0166 0.01637 0.01584 0.0156 0.01494 0.01437 0.01352 0.01333 0.01304
Cumulative Proportion  0.75166 0.7707 0.78901 0.80727 0.82468 0.8413 0.85764 0.87349 0.8891 0.90403 0.91840 0.93192 0.94525 0.95829
                          PC29    PC30    PC31    PC32
Standard deviation     0.61380 0.60237 0.56256 0.52784
Proportion of Variance 0.01177 0.01134 0.00989 0.00871
Cumulative Proportion  0.97006 0.98140 0.99129 1.00000

#We can lot this to get an idea of which components we should keep and which we should drop

plot(pca8, type = "lines")
we should drop after pc6 since there were not much variance is shown after pc6
```

```{r}
Now print out the eigenvectors (often called loadings) for the components you generated:

pca8$rotation

   PC1         PC2         PC3         PC4          PC5          PC6           PC7          PC8          PC9        PC10
Q1_AF   0.21939788  0.14425900 -0.09372746  0.23505039 -0.025950069  0.023284939 -0.1012021091  0.049003689 -0.061534083 -0.06935476
Q2_SE  -0.20898123 -0.13035347 -0.09511445  0.21385731 -0.211603645  0.026807178  0.1045267133 -0.144260993 -0.036157931 -0.07763362
Q3_AG  -0.17399032  0.15740465  0.20639039  0.08832221  0.110888964  0.187496407  0.1677039712  0.124205526 -0.203071109  0.25053515
Q4_SD  -0.14736521  0.26364779 -0.18831073 -0.09445513 -0.026573433  0.116547736 -0.1555747724 -0.202791860 -0.178516399  0.22449268
Q5_AF  -0.23161793 -0.13540153  0.05119129 -0.14919685  0.180571839 -0.292528699 -0.0792117582  0.111988581 -0.256634452  0.03424689
Q6_SE  -0.20555790 -0.14371972 -0.04102976  0.11354302  0.110528587  0.191774729 -0.4820700652  0.215911581 -0.007484659 -0.06419635
Q7_AG   0.05921923 -0.21667960 -0.26300949 -0.16030490 -0.004885512  0.087708700  0.1643137424  0.093205720 -0.503820929  0.10938538
Q8_SD  -0.17796928  0.27534997 -0.19611487 -0.12306184 -0.062570040 -0.017244228 -0.0349900765  0.069850280 -0.179329221 -0.26650293
Q9_AF   0.18082834  0.07507400 -0.05245775  0.20724432  0.109326864  0.189808878  0.0841277571 -0.068980794 -0.281697581 -0.38902991
Q10_SE -0.22377869 -0.09068557 -0.16192345  0.26932733 -0.174957842 -0.107059340  0.1473586406 -0.070519578  0.159497504  0.01038556
Q11_AG -0.09016370  0.16091695  0.21581509  0.24692689  0.192167605 -0.239601160 -0.1909269458 -0.239793065 -0.316600924 -0.14445755
Q12_SD -0.17455027  0.21931490 -0.17568194 -0.19539636 -0.038535384 -0.019407929  0.0245561737  0.176126722  0.130175473 -0.12230097
Q13_AF -0.21809742 -0.16368269  0.04518596 -0.15246661  0.148729177  0.345888755 -0.0008269885 -0.265087376 -0.139077323 -0.26397271
Q14_SE -0.23089006 -0.16732041 -0.11545749  0.18858516 -0.113935836  0.068735304  0.0262780678  0.045967735 -0.159974713 -0.05516617
Q15_AG  0.11802365 -0.19007207 -0.33002097 -0.06463064  0.161695105 -0.102262963 -0.0303895892 -0.239388141  0.008771057 -0.13889066
Q16_SD  0.14656689 -0.23137984  0.03069187  0.18823214  0.323018417  0.023647032  0.1859782792 -0.197205038 -0.173666620  0.16579852
Q17_AF  0.22808186  0.15032483 -0.12806301  0.26217769 -0.012688613  0.081426042 -0.0613150468  0.005526549 -0.099610148 -0.08952747
Q18_SE -0.19844699 -0.10152645 -0.17450512  0.31787288 -0.119997529 -0.026714420  0.1970885893 -0.170639002  0.201729666 -0.04732850
Q19_AG -0.20716496  0.10561539  0.12665504  0.12698531  0.291748441 -0.125000089 -0.1217721847 -0.230370948  0.120614361 -0.06592467
Q20_SD -0.15575377  0.33432360 -0.16961349 -0.08848565 -0.024422742 -0.056644454 -0.0101948501  0.076588631 -0.005871440 -0.26229743
Q21_AF -0.21372343 -0.15421689 -0.06276763 -0.14705247  0.292049775  0.063033585  0.1774961522  0.055023420  0.077013726  0.15786880
Q22_SE  0.12946989  0.08961723 -0.04506464 -0.08361504  0.478612886  0.244243238  0.0564848538  0.099294533  0.322982980 -0.08459737
Q23_AG  0.11391601 -0.08802318 -0.28725962 -0.04816672  0.246836745 -0.410012310 -0.1772882231 -0.061543218  0.055611249 -0.13585859
Q24_SD -0.07506380  0.24240448 -0.22657742  0.01872029  0.163707022  0.011758431 -0.0963854832 -0.357134800  0.052236264  0.45745367
Q25_AF  0.20451431  0.19774580 -0.10998688  0.24022477  0.047668480 -0.321527741  0.0594434570  0.220912650 -0.077777370  0.22710006
Q26_SE -0.22483686 -0.13070433 -0.15841857  0.21940804  0.006297694 -0.098814025  0.0751597950  0.133956707  0.071648614  0.04294618
Q27_AG -0.10836500  0.17253321  0.21223553  0.18620037  0.226455511 -0.058341046  0.3329259806  0.301431910 -0.118975038 -0.06351345
Q28_SD -0.19086182  0.10052681 -0.05054739  0.07389942  0.242550393 -0.034348851  0.2591527500  0.054326716  0.180061631 -0.15589728
Q29_AF  0.15021196  0.17237752 -0.23157405  0.17585773  0.019902419  0.408672680  0.0679736701 -0.051028728  0.080564399  0.05681054
Q30_SE -0.12536356 -0.12855220 -0.10425608  0.26757330  0.121725439  0.185073577 -0.4521164585  0.351161161  0.019571635  0.15527818
Q31_AG  0.10756558 -0.14151134 -0.37671968 -0.07039664  0.126078093 -0.043393865  0.1407480829  0.247404189 -0.065020385 -0.02280943
Q32_SD -0.19346976  0.23951877 -0.19451885 -0.12150291 -0.017973572  0.008655502  0.0926786337 -0.026186771 -0.150369855  0.16767592
              PC11          PC12          PC13        PC14         PC15         PC16        PC17          PC18         PC19
Q1_AF   0.14061884  0.0322951228  0.3756185888 -0.13660090 -0.001985660 -0.111980588  0.15825413  0.4790918710  0.099576303
Q2_SE   0.19241882 -0.1509661279  0.1245183868 -0.22360874 -0.082415369 -0.108822683 -0.35991180  0.2149584843  0.064868453
Q3_AG  -0.03533538  0.2580817344  0.1767842090 -0.22471826  0.468713506 -0.172327078  0.09585658 -0.0329087165  0.161532902
Q4_SD   0.10841711  0.1395494240  0.0594485555  0.10344737  0.167072429  0.046394950 -0.26615140 -0.2740685429 -0.004542307
Q5_AF  -0.01674378  0.1659454516 -0.0070361958 -0.05170257 -0.142255145 -0.248173358 -0.13596612  0.2849603067  0.007819186
Q6_SE  -0.00303950  0.3053188241 -0.0318869791  0.04569763 -0.062933217 -0.091267469 -0.03523054  0.0007054778 -0.275548680
Q7_AG  -0.23597045  0.0056448709 -0.0006248341  0.07702940 -0.338112878  0.274056262 -0.03379258 -0.0233963814  0.340668817
Q8_SD  -0.11474553 -0.0338721548  0.1527114626  0.08583795 -0.136861923 -0.238174255  0.21645101 -0.2287198224 -0.115593250
Q9_AF   0.13695639  0.0003245939 -0.6291815703  0.16958690  0.153169979 -0.314957090 -0.13543606  0.0389485422  0.111742286
Q10_SE -0.17852414  0.1982641437 -0.0538450754 -0.02777156 -0.116685629 -0.119166764 -0.09455533 -0.0984932847 -0.028443543
Q11_AG -0.05497252 -0.1367546739  0.1798572526  0.01640181  0.002719618  0.170678737 -0.29188229 -0.0895085696 -0.071728702
Q12_SD -0.02324821  0.1564043449 -0.0530243909  0.04061232  0.058740868  0.284278359 -0.40778981  0.3705563155 -0.155360293
Q13_AF  0.13665830 -0.1000471540  0.0814799832 -0.23547707 -0.037941636  0.088503021  0.09959034  0.1281148622 -0.039428383
Q14_SE -0.14257534  0.0037910786 -0.0205944294 -0.17639105  0.353560798  0.190125613  0.08869389  0.0049015079  0.051570500
Q15_AG -0.02451440  0.2605762719  0.1175277075  0.09220484  0.018979537  0.036297157  0.18684975  0.1747048402 -0.014109125
Q16_SD -0.09566452  0.1275646566  0.2040699578  0.20258398 -0.013022256 -0.052977495 -0.04768739 -0.1418116897 -0.393660155
Q17_AF -0.10669286  0.2445831398 -0.0088982831 -0.19533787 -0.105044843  0.108811672  0.27384226  0.0517806111  0.034984269
Q18_SE -0.14707670 -0.1454201119  0.0756638623 -0.08913721 -0.102695257 -0.186832566 -0.01847774 -0.1524923775  0.073458414
Q19_AG -0.31939208 -0.1241597187 -0.0205263300  0.22912012  0.011237074  0.137988707  0.06888718  0.1649243515  0.351224702
Q20_SD -0.23939206 -0.1167657635  0.1584998499  0.13722374  0.015859123 -0.195914131  0.16196508 -0.0305535375 -0.096208359
Q21_AF  0.03107467 -0.3148811388  0.0147993316  0.27954631  0.199723183 -0.225807636  0.15023776  0.2547411407 -0.008424361
Q22_SE -0.38461663  0.0918192791 -0.0193005076 -0.36294636 -0.075139514 -0.058551327 -0.27704711 -0.1015507454  0.157760253
Q23_AG  0.19187347 -0.1869485406 -0.0634938733 -0.30436777  0.317678711  0.185567995  0.08042081 -0.1932780163 -0.028788108
Q24_SD  0.13670696  0.0152466936 -0.2711181731 -0.19288609 -0.278960099 -0.156988607  0.11688211  0.1345499021 -0.129977800
Q25_AF -0.04744618 -0.0220911434 -0.0009406745  0.11588887  0.062623040 -0.090463447 -0.14862654  0.1168786176  0.132315275
Q26_SE -0.11310328  0.1639465348 -0.2672909716  0.11627237  0.164501940  0.252361815  0.17302696  0.0856197928 -0.168367715
Q27_AG  0.10115775 -0.1811401042 -0.1020104548 -0.19909177 -0.282746628  0.207605249  0.12472104  0.0077198003 -0.350495051
Q28_SD  0.55636856  0.3413988594  0.1568063558  0.22388120 -0.148025938  0.132299128  0.03981873 -0.1950467243  0.324236006
Q29_AF  0.00106400 -0.2176372943  0.1485027202  0.26034358  0.094610200  0.254374776 -0.10474092  0.1111712096 -0.186622909
Q30_SE  0.13692683 -0.2906998236  0.0057937572  0.06349187 -0.156501115  0.008003983  0.03154537 -0.1374980827  0.216931430
Q31_AG  0.08332603 -0.1294672484  0.1577515083 -0.10213417  0.086943654 -0.199377253 -0.20056007 -0.1229690196 -0.100460362
Q32_SD  0.02939022 -0.1154703733 -0.1577800029 -0.16858544  0.027666837  0.126109795  0.12963875 -0.0133449257  0.068349247
               PC20         PC21         PC22          PC23         PC24         PC25         PC26         PC27         PC28
Q1_AF   0.087695779  0.129795561 -0.173983935 -0.3530898893 -0.038537038 -0.142583797  0.011365828  0.263787606  0.154074473
Q2_SE  -0.277484256 -0.217723192  0.553788920  0.0007478772 -0.064802396  0.053324564 -0.081717455 -0.055696433 -0.078418191
Q3_AG   0.065408289 -0.375992316 -0.022892923  0.2020056650  0.026143304 -0.134766603 -0.191797668 -0.002453889  0.032361258
Q4_SD  -0.400085780  0.217271047 -0.065494323 -0.3320132455  0.220088146 -0.225938936  0.089415791  0.089847951 -0.175962980
Q5_AF  -0.101859810  0.330523548 -0.151432759  0.1125701750 -0.022575806  0.044745984 -0.106915827 -0.380043586 -0.010204308
Q6_SE  -0.136256648  0.021403975  0.106214880  0.1307162114 -0.409800985 -0.103833069  0.076350655  0.277530990  0.198315315
Q7_AG   0.008640330 -0.096669871  0.074761098  0.0216744350  0.029318067 -0.087257951 -0.078387767  0.258906284  0.262890408
Q8_SD   0.084104008 -0.011340402  0.184643991  0.0495941692  0.082446331 -0.023370038 -0.031784785 -0.154032719 -0.007345520
Q9_AF   0.011707810 -0.079148407 -0.091524273 -0.0506014153 -0.005743351 -0.013946946  0.020706244  0.021902464  0.042906535
Q10_SE  0.088282634  0.014164475 -0.263878898 -0.0670586693  0.174905125 -0.327984143 -0.096391846 -0.122394831  0.261680957
Q11_AG  0.520000301  0.030027393  0.116693193 -0.0654257107 -0.011503083 -0.057555793  0.063096675  0.050298791 -0.052073370
Q12_SD  0.064882464 -0.327224303 -0.257813621 -0.0755235994  0.149284102  0.280285907 -0.010965030  0.044891879  0.039739801
Q13_AF  0.007833270  0.165485746 -0.204471157  0.0665394588  0.178315690 -0.040839801 -0.253964039 -0.135825478  0.025744631
Q14_SE  0.020425515  0.268311138 -0.005381689  0.1794445244  0.141630610  0.307416972  0.570896014  0.020462671  0.077501880
Q15_AG  0.028858507 -0.364220539  0.039140220  0.0485229324  0.002495994 -0.329457951  0.391541853 -0.262124663 -0.267986187
Q16_SD -0.133741432 -0.170170918 -0.093145128 -0.1998952880 -0.057228214  0.405487012 -0.076593183 -0.076078661  0.116422772
Q17_AF -0.192809293  0.123782948  0.084577801 -0.0869027648  0.112741038  0.330038872 -0.184942396 -0.086109890 -0.256188872
Q18_SE -0.020207079 -0.003892791 -0.285106366 -0.0556695320 -0.171903043  0.066484044  0.040130725  0.096192126  0.003227829
Q19_AG -0.316749051 -0.048669709 -0.144939233  0.2025006296 -0.120848627 -0.007642626 -0.120082145  0.208205321 -0.226603695
Q20_SD -0.087382510 -0.108210687  0.080740120  0.0313975317  0.095103212  0.169502013 -0.047675757  0.080999940  0.165259791
Q21_AF  0.023524515  0.104300420  0.194974526 -0.2861491608  0.109840844 -0.006610400  0.113823236  0.088078819  0.091155153
Q22_SE  0.076637599  0.130169226  0.218538004 -0.1758712081  0.008363693 -0.047225578  0.099540000 -0.097299313  0.053627606
Q23_AG -0.194202501 -0.089053550  0.001641600  0.0194345091  0.022005197 -0.043700198 -0.208234678  0.019374350  0.373477542
Q24_SD  0.198765464 -0.035395750  0.064588183  0.2720535724  0.183608546  0.128682015  0.062663583  0.206243952  0.060234556
Q25_AF -0.114351745  0.138192675  0.078663981  0.1644645263  0.026844922 -0.025212735  0.058669263 -0.207903777  0.117388113
Q26_SE  0.221813378  0.143140848  0.286215694 -0.1934578547  0.071630154 -0.121216150 -0.386701131  0.054534001 -0.213510130
Q27_AG -0.282362871 -0.084299473 -0.077299886  0.0029983152  0.084929782 -0.284042666  0.230063138  0.075006183 -0.081684497
Q28_SD  0.083280254  0.121117270  0.068542416  0.0735653510 -0.050069972  0.156612740  0.040771059  0.027354784  0.092755248
Q29_AF  0.005873596  0.183535288  0.001172524  0.3616599773 -0.130993748 -0.172263466 -0.122501077 -0.279610304  0.148365344
Q30_SE  0.079788414 -0.279279597 -0.149438603 -0.1331504797  0.229488893  0.095887084 -0.000547371 -0.272635445 -0.102041697
Q31_AG  0.140928986  0.108659823 -0.188819499  0.2380857806 -0.076859656  0.012696967 -0.117691257  0.310319849 -0.504205249
Q32_SD  0.093458663 -0.035893159 -0.107321503 -0.2840141530 -0.676287614  0.073831272  0.038661291 -0.256031926 -0.054422767
                PC29         PC30         PC31         PC32
Q1_AF  -0.2948854986  0.064010874  0.017327812 -0.029396635
Q2_SE  -0.1070933760 -0.039487402  0.035337562  0.088014191
Q3_AG   0.0759656114  0.109461704 -0.051382244 -0.011354488
Q4_SD   0.0168385166  0.073713373  0.047213912 -0.068889415
Q5_AF  -0.0239150230  0.403904603  0.055332373  0.080712344
Q6_SE   0.1784002545 -0.133011623 -0.066924282 -0.047450645
Q7_AG   0.0613755309  0.077416335 -0.008757848 -0.023885051
Q8_SD  -0.3333967450 -0.050187942 -0.536883777 -0.120886878
Q9_AF  -0.0527600236  0.060968974  0.012321628  0.009468763
Q10_SE -0.0153453043 -0.309966587 -0.001295353  0.466009559
Q11_AG  0.2391854633  0.014007008 -0.016697567  0.087451346
Q12_SD  0.0591502171 -0.011016448 -0.220197111 -0.073527958
Q13_AF  0.1098357698 -0.401980523  0.104055804 -0.310048464
Q14_SE -0.2024719528 -0.019194474  0.024044376  0.059835045
Q15_AG  0.1022314982  0.029445673  0.090647001 -0.072344173
Q16_SD -0.2571768330 -0.086716335  0.037674859 -0.026097677
Q17_AF  0.4383960746 -0.016941813 -0.198550185  0.269304217
Q18_SE  0.3021504176  0.353796443 -0.148002010 -0.440328309
Q19_AG -0.2418213580 -0.116303542 -0.097837819  0.106490149
Q20_SD  0.1459326526  0.048640941  0.656929581  0.053105884
Q21_AF  0.3393529445 -0.069351571 -0.252591963  0.192860856
Q22_SE -0.1099775312  0.033285399  0.042889279 -0.048373326
Q23_AG  0.0034455552  0.113283998 -0.112699012  0.054746003
Q24_SD -0.0761592438  0.035727620  0.028959320 -0.013166693
Q25_AF  0.1441659674 -0.471479028  0.009130771 -0.426318974
Q26_SE -0.1739700129  0.107101680  0.151477836 -0.257127083
Q27_AG  0.0029212178  0.057499682  0.012101012  0.014221321
Q28_SD -0.0023911172  0.008447881  0.053721908  0.029872765
Q29_AF -0.0006825951  0.253576994 -0.037177138  0.132536501
Q30_SE -0.0705213588  0.031187516  0.090241678  0.007308784
Q31_AG -0.0357714645 -0.146588546  0.055600970  0.160962227
Q32_SD -0.0343270163 -0.178452859  0.086274365  0.104239206

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadingsH <- abs(pca8$rotation) #abs() will make all eigenvectors positive

sweep(loadingsH, 2, colSums(loadingsH), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

  PC1        PC2         PC3         PC4         PC5         PC6         PC7         PC8         PC9        PC10
Q1_AF  0.04037645 0.02714296 0.019046867 0.045740859 0.005851656 0.005521076 0.023220761 0.010301782 0.013687415 0.015145482
Q2_SE  0.03845945 0.02452658 0.019328725 0.041616681 0.047715935 0.006356232 0.023983589 0.030327214 0.008042837 0.016953396
Q3_AG  0.03201996 0.02961638 0.041941713 0.017187522 0.025005101 0.044457146 0.038479571 0.026111061 0.045170390 0.054711107
Q4_SD  0.02712006 0.04960649 0.038267647 0.018380988 0.005992223 0.027634555 0.035696534 0.042631843 0.039708531 0.049024033
Q5_AF  0.04262535 0.02547639 0.010402861 0.029033740 0.040718363 0.069361282 0.018175088 0.023542758 0.057084823 0.007478732
Q6_SE  0.03782944 0.02704150 0.008337880 0.022095496 0.024923837 0.045471577 0.110610675 0.045389932 0.001664860 0.014019004
Q7_AG  0.01089829 0.04076922 0.053447588 0.031195368 0.001101667 0.020796551 0.037701685 0.019594138 0.112068072 0.023887247
Q8_SD  0.03275222 0.05180831 0.039853570 0.023947861 0.014109341 0.004088767 0.008028451 0.014684249 0.039889332 0.058198105
Q9_AF  0.03327839 0.01412550 0.010660225 0.040329791 0.024652853 0.045005455 0.019303061 0.014501462 0.062659773 0.084955175
Q10_SE 0.04118267 0.01706289 0.032905345 0.052411158 0.039452425 0.025384768 0.033811348 0.014824952 0.035478038 0.002267967
Q11_AG 0.01659310 0.03027723 0.043856959 0.048052027 0.043333171 0.056811669 0.043808068 0.050410408 0.070423544 0.031546203
Q12_SD 0.03212301 0.04126506 0.035701283 0.038024175 0.008689604 0.004601801 0.005634399 0.037026174 0.028955753 0.026707715
Q13_AF 0.04013713 0.03079762 0.009182485 0.029670035 0.033537946 0.082013449 0.000189752 0.055727895 0.030935848 0.057645563
Q14_SE 0.04249139 0.03148207 0.023462745 0.036698715 0.025692161 0.016297782 0.006029486 0.009663550 0.035584186 0.012047021
Q15_AG 0.02172025 0.03576290 0.067065357 0.012577137 0.036461721 0.024247502 0.006972872 0.050325283 0.001951002 0.030330521
Q16_SD 0.02697315 0.04353513 0.006237062 0.036630016 0.072839604 0.005606932 0.042672600 0.041457356 0.038629763 0.036206579
Q17_AF 0.04197459 0.02828428 0.026024381 0.051019836 0.002861241 0.019306874 0.014068699 0.001161817 0.022156915 0.019550738
Q18_SE 0.03652080 0.01910265 0.035462134 0.061858134 0.027059053 0.006334238 0.045221854 0.035872521 0.044872004 0.010335455
Q19_AG 0.03812519 0.01987200 0.025738261 0.024711370 0.065788326 0.029638687 0.027940552 0.048429647 0.026829014 0.014396430
Q20_SD 0.02866384 0.06290445 0.034468080 0.017219327 0.005507249 0.013430928 0.002339202 0.016100816 0.001306021 0.057279720
Q21_AF 0.03933217 0.02901659 0.012755351 0.028616442 0.065856276 0.014945851 0.040726381 0.011567278 0.017130650 0.034474911
Q22_SE 0.02382673 0.01686188 0.009157830 0.016271505 0.107925651 0.057912349 0.012960414 0.020874156 0.071843145 0.018474117
Q23_AG 0.02096431 0.01656195 0.058375589 0.009373253 0.055660884 0.097217742 0.040678672 0.012937900 0.012369961 0.029668388
Q24_SD 0.01381422 0.04560947 0.046044030 0.003642973 0.036915402 0.002788034 0.022115589 0.075078530 0.011619242 0.099897349
Q25_AF 0.03763739 0.03720674 0.022351031 0.046747793 0.010749087 0.076237225 0.013639264 0.046441280 0.017300512 0.049593425
Q26_SE 0.04137741 0.02459259 0.032193098 0.042696853 0.001420109 0.023429727 0.017245368 0.028160999 0.015937254 0.009378457
Q27_AG 0.01994274 0.03246288 0.043129537 0.036234633 0.051064982 0.013833206 0.076389658 0.063368411 0.026464369 0.013869875
Q28_SD 0.03512488 0.01891456 0.010272009 0.014380843 0.054694326 0.008144433 0.059462436 0.011420814 0.040052246 0.034044375
Q29_AF 0.02764396 0.03243359 0.047059422 0.034221954 0.004487931 0.096900103 0.015596516 0.010727495 0.017920448 0.012406115
Q30_SE 0.02307103 0.02418766 0.021186445 0.052069823 0.027448691 0.043882671 0.103737839 0.073822724 0.004353442 0.033909180
Q31_AG 0.01979562 0.02662598 0.076555256 0.013699202 0.028430200 0.010289090 0.032294560 0.052010453 0.014462895 0.004981053
Q32_SD 0.03560482 0.04506651 0.039529233 0.023644492 0.004052982 0.002052300 0.021265055 0.005505104 0.033447717 0.036616561
               PC11         PC12         PC13        PC14         PC15        PC16        PC17        PC18        PC19        PC20
Q1_AF  0.0329626017 6.760437e-03 0.0963770502 0.027596548 0.0004633842 0.022064102 0.033284864 0.107420471 0.022813501 0.021033175
Q2_SE  0.0451050875 3.160220e-02 0.0319492037 0.045174149 0.0192328856 0.021441884 0.075698594 0.048197315 0.014861734 0.066552519
Q3_AG  0.0082830027 5.402504e-02 0.0453596842 0.045398297 0.1093814583 0.033954476 0.020161074 0.007378689 0.037008112 0.015687688
Q4_SD  0.0254141622 2.921231e-02 0.0152534422 0.020898766 0.0389889042 0.009141432 0.055978399 0.061450786 0.001040669 0.095957576
Q5_AF  0.0039249254 3.473787e-02 0.0018053627 0.010445117 0.0331974116 0.048898852 0.028597130 0.063892903 0.001791420 0.024430312
Q6_SE  0.0007124922 6.391332e-02 0.0081816318 0.009231981 0.0146864278 0.017982891 0.007409878 0.000158180 0.063129779 0.032680136
Q7_AG  0.0553140679 1.181658e-03 0.0001603213 0.015561724 0.0789038062 0.053998691 0.007107437 0.005245863 0.078049175 0.002072318
Q8_SD  0.0268976141 7.090562e-03 0.0391830456 0.017341254 0.0319388210 0.046928678 0.045525146 0.051282839 0.026483075 0.020171716
Q9_AF  0.0321040825 6.794824e-05 0.1614368021 0.034260484 0.0357445550 0.062057588 0.028485644 0.008732920 0.025600797 0.002808031
Q10_SE 0.0418480213 4.150324e-02 0.0138156888 0.005610499 0.0272303745 0.023480030 0.019887388 0.022083854 0.006516579 0.021173928
Q11_AG 0.0128861638 2.862727e-02 0.0461481726 0.003313546 0.0006346645 0.033629695 0.061390260 0.020069330 0.016433456 0.124718175
Q12_SD 0.0054496366 3.274060e-02 0.0136051158 0.008204630 0.0137080789 0.056012803 0.085768556 0.083084971 0.035593932 0.015561573
Q13_AF 0.0320342070 2.094318e-02 0.0209063147 0.047571826 0.0088542606 0.017438198 0.020946377 0.028725511 0.009033268 0.001878751
Q14_SE 0.0334212269 7.935981e-04 0.0052841644 0.035635079 0.0825088143 0.037461411 0.018654579 0.001099001 0.011815097 0.004898907
Q15_AG 0.0057464443 5.454723e-02 0.0301555197 0.018627514 0.0044291648 0.007151812 0.039299248 0.039171769 0.003232481 0.006921497
Q16_SD 0.0224248142 2.670350e-02 0.0523607222 0.040926659 0.0030389424 0.010438424 0.010029869 0.031796571 0.090189794 0.032076880
Q17_AF 0.0250099783 5.119934e-02 0.0022831412 0.039462777 0.0245138192 0.021439714 0.057595984 0.011610086 0.008015096 0.046243864
Q18_SE 0.0344763954 3.044124e-02 0.0194140015 0.018007781 0.0239655074 0.036812565 0.003886338 0.034191361 0.016829743 0.004846516
Q19_AG 0.0748690150 2.599073e-02 0.0052666913 0.046287574 0.0026223429 0.027188613 0.014488724 0.036978819 0.080467589 0.075969886
Q20_SD 0.0561161320 2.444293e-02 0.0406682429 0.027722377 0.0037009687 0.038601952 0.034065370 0.006850618 0.022041886 0.020958040
Q21_AF 0.0072842444 6.591503e-02 0.0037972453 0.056474834 0.0466084563 0.044492021 0.031598817 0.057117257 0.001930069 0.005642179
Q22_SE 0.0901583672 1.922081e-02 0.0049521670 0.073323577 0.0175349537 0.011536664 0.058270045 0.022769388 0.036143777 0.018380954
Q23_AG 0.0449772511 3.913451e-02 0.0162913988 0.061489344 0.0741351810 0.036563400 0.016914540 0.043336188 0.006595520 0.046578014
Q24_SD 0.0320456152 3.191637e-03 0.0695641019 0.038967461 0.0650996013 0.030932259 0.024583276 0.030168355 0.029778658 0.047672407
Q25_AF 0.0111219069 4.624407e-03 0.0002413604 0.023412238 0.0146140432 0.017824470 0.031259937 0.026206156 0.030314186 0.027426409
Q26_SE 0.0265126529 3.431943e-02 0.0685821100 0.023489714 0.0383890411 0.049724125 0.036391965 0.019197400 0.038574007 0.053200276
Q27_AG 0.0237124877 3.791861e-02 0.0261740686 0.040221152 0.0659832459 0.040905512 0.026232003 0.001730909 0.080300422 0.067722618
Q28_SD 0.1304189068 7.146607e-02 0.0402337224 0.045229192 0.0345441143 0.026067566 0.008374889 0.043732762 0.074284324 0.019974145
Q29_AF 0.0002494133 4.555868e-02 0.0381031572 0.052595438 0.0220787357 0.050120749 0.022029675 0.024926459 0.042756376 0.001408738
Q30_SE 0.0320971549 6.085308e-02 0.0014865751 0.012826830 0.0365219264 0.001577065 0.006634793 0.030829387 0.049700232 0.019136653
Q31_AG 0.0195325375 2.710177e-02 0.0404762317 0.020633470 0.0202896302 0.039284308 0.042182877 0.027571726 0.023016044 0.033800761
Q32_SD 0.0068893911 2.417177e-02 0.0404835429 0.034058165 0.0064564791 0.024848051 0.027266322 0.002992157 0.015659204 0.022415360
               PC21         PC22         PC23         PC24        PC25         PC26         PC27         PC28         PC29
Q1_AF  0.0284822647 0.0392046351 0.0778279839 0.0101779203 0.033002115 0.0028240968 0.0575681598 0.0354375315 0.0693519561
Q2_SE  0.0477770543 0.1247879152 0.0001648469 0.0171147979 0.012342380 0.0203045477 0.0121550106 0.0180363890 0.0251865051
Q3_AG  0.0825075415 0.0051585723 0.0445260375 0.0069046422 0.031192765 0.0476564635 0.0005355289 0.0074431740 0.0178657946
Q4_SD  0.0476778358 0.0147581500 0.0731822754 0.0581269273 0.052295301 0.0222173734 0.0196081282 0.0404719454 0.0039601271
Q5_AF  0.0725299007 0.0341230705 0.0248126894 0.0059624395 0.010356802 0.0265656526 0.0829394915 0.0023470175 0.0056243987
Q6_SE  0.0046968763 0.0239339087 0.0288124342 0.1082315087 0.024032961 0.0189710451 0.0605674718 0.0456130408 0.0419566465
Q7_AG  0.0212131820 0.0168462770 0.0047774735 0.0077431210 0.020196523 0.0194772116 0.0565028759 0.0604654810 0.0144344606
Q8_SD  0.0024885313 0.0416067167 0.0109315342 0.0217746932 0.005409175 0.0078976480 0.0336156058 0.0016894888 0.0784091335
Q9_AF  0.0173682817 0.0206236036 0.0111535511 0.0015168620 0.003228128 0.0051449340 0.0047799235 0.0098686151 0.0124082427
Q10_SE 0.0031082444 0.0594610986 0.0147810549 0.0461937531 0.075914448 0.0239507317 0.0267110548 0.0601873040 0.0036089495
Q11_AG 0.0065891942 0.0262950372 0.0144211186 0.0030380503 0.013321730 0.0156777943 0.0109770466 0.0119770113 0.0562522737
Q12_SD 0.0718059163 0.0580943807 0.0166468926 0.0394270490 0.064874325 0.0027245094 0.0097970596 0.0091402582 0.0139111054
Q13_AF 0.0363140987 0.0460744672 0.0146666106 0.0470945088 0.009452686 0.0631031029 0.0296421161 0.0059213324 0.0258314686
Q14_SE 0.0588780447 0.0012126818 0.0395531166 0.0374057046 0.071154018 0.1418520121 0.0044657075 0.0178256349 0.0476178926
Q15_AG 0.0799243495 0.0088196537 0.0106954125 0.0006592108 0.076255572 0.0972874189 0.0572052446 0.0616375236 0.0240430758
Q16_SD 0.0373422102 0.0209888386 0.0440608687 0.0151143999 0.093853082 0.0190313066 0.0166031626 0.0267775418 0.0604835319
Q17_AF 0.0271628601 0.0190583218 0.0191550854 0.0297757523 0.076390031 0.0459531164 0.0187923457 0.0589241101 0.1031031553
Q18_SE 0.0008542318 0.0642443859 0.0122706641 0.0454008808 0.015388242 0.0099713852 0.0209926606 0.0007424092 0.0710605391
Q19_AG 0.0106800535 0.0326598530 0.0446351374 0.0319170273 0.001768945 0.0298371218 0.0454380605 0.0521194422 0.0568721904
Q20_SD 0.0237456922 0.0181935588 0.0069206360 0.0251174704 0.039232542 0.0118461190 0.0176771668 0.0380101840 0.0343208295
Q21_AF 0.0228876253 0.0439345456 0.0630729255 0.0290097895 0.001530028 0.0282819544 0.0192220385 0.0209659235 0.0798099284
Q22_SE 0.0285642614 0.0492442174 0.0387654871 0.0022089138 0.010930723 0.0247329616 0.0212342895 0.0123344895 0.0258648084
Q23_AG 0.0195418607 0.0003699095 0.0042837496 0.0058117372 0.010114746 0.0517406101 0.0042281960 0.0859008109 0.0008103348
Q24_SD 0.0077672233 0.0145539652 0.0599659795 0.0484923918 0.029784440 0.0155701828 0.0450100176 0.0138541054 0.0179113336
Q25_AF 0.0303249227 0.0177257325 0.0362512292 0.0070899448 0.005835681 0.0145777037 0.0453722525 0.0269995729 0.0339053358
Q26_SE 0.0314107470 0.0644943560 0.0426419313 0.0189180600 0.028056408 0.0960846321 0.0119013252 0.0491078881 0.0409147305
Q27_AG 0.0184986287 0.0174183544 0.0006608879 0.0224305913 0.065743856 0.0571643841 0.0163691085 0.0187876479 0.0006870198
Q28_SD 0.0265779054 0.0154449917 0.0162152560 0.0132238545 0.036249221 0.0101304907 0.0059698201 0.0213339494 0.0005623493
Q29_AF 0.0402748800 0.0002642105 0.0797170005 0.0345964296 0.039871702 0.0304381601 0.0610212547 0.0341244170 0.0001605345
Q30_SE 0.0612849571 0.0336737175 0.0293489950 0.0606097350 0.022193802 0.0001360067 0.0594990838 0.0234698570 0.0165854008
Q31_AG 0.0238442502 0.0425476040 0.0524788074 0.0202992105 0.002938811 0.0292430517 0.0677232071 0.1159685251 0.0084128282
Q32_SD 0.0078763745 0.0241832693 0.0626023276 0.1786126225 0.017088815 0.0096062711 0.0558755851 0.0125173787 0.0080731190
              PC30         PC31        PC32
Q1_AF  0.016303704 0.0051566596 0.007564102
Q2_SE  0.010057524 0.0105162602 0.022647094
Q3_AG  0.027880125 0.0152910677 0.002921644
Q4_SD  0.018774950 0.0140505956 0.017726062
Q5_AF  0.102875348 0.0164666041 0.020768242
Q6_SE  0.033878339 0.0199162911 0.012209613
Q7_AG  0.019718103 0.0026062865 0.006145907
Q8_SD  0.012782974 0.1597736021 0.031105625
Q9_AF  0.015528925 0.0036668475 0.002436425
Q10_SE 0.078949139 0.0003854898 0.119909778
Q11_AG 0.003567614 0.0049691023 0.022502267
Q12_SD 0.002805912 0.0655294259 0.018919614
Q13_AF 0.102385281 0.0309664238 0.079779141
Q14_SE 0.004888873 0.0071554716 0.015396266
Q15_AG 0.007499875 0.0269760394 0.018615012
Q16_SD 0.022086832 0.0112118268 0.006715241
Q17_AF 0.004315115 0.0590874218 0.069295164
Q18_SE 0.090112695 0.0440445684 0.113301688
Q19_AG 0.029622756 0.0291159863 0.027401176
Q20_SD 0.012388950 0.1954985603 0.013664773
Q21_AF 0.017663990 0.0751699521 0.049625382
Q22_SE 0.008477861 0.0127636091 0.012447030
Q23_AG 0.028853672 0.0335385941 0.014086795
Q24_SD 0.009099900 0.0086181312 0.003387946
Q25_AF 0.120086696 0.0027172663 0.109696921
Q26_SE 0.027279022 0.0450789548 0.066161843
Q27_AG 0.014645289 0.0036011934 0.003659314
Q28_SD 0.002151693 0.0159873387 0.007686616
Q29_AF 0.064586591 0.0110637077 0.034103211
Q30_SE 0.007943526 0.0268554176 0.001880637
Q31_AG 0.037336410 0.0165465369 0.041417487
Q32_SD 0.045452317 0.0256747673 0.026821982

#Now examine your components and try to come up with substantive descriptions of what some might represent?

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca8)

By loading and sweeping, there were not many significant finding as the class data, but after running a biplot we can clearly see that certain arrows are headed to certain points. As authors divided into four measures of their questionnaire (Self defeating, self enhancing, aggressive, and affiliative), it is clearly shown in the data that each direction means different measures.It is intersting how ag and sd are in the end and af and se is in the end. The HSQ serves a great tool for checking individual's humor style on daily life basis. Based on the result, it might be able to explain if depressed participant's joke as a indicator if participant is getting better. (To decide whether theier joke isas a agrressive (sarcasm) or genuinely self enhancing).
```


