---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session
```{r}
install.packages("corrplot")
install.packages("FactoMineR")
install.packages("factoextra")
library(FactoMineR)
library(factoextra)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv")
D1 <- D1[,-1]
  #We won't need the id variable, so remove that.


```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

#Correlogram 1: Assistments Variables
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank", main = "Assistments Variables")

#Study your correlogram image and save it, you will need it later
```

## Create a new data frame with the mean_correct variables removed

```{r}
D1.1 <- D1[,-4]

#The, scale and center your data for easier interpretation
D1.1 <- scale(D1.1, center = TRUE)
```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D1.1, scale = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

#PCA Variance Plot 1: Assistments
plot(pca, type = "lines", main = "Assistments PCs")
```

## Q1:Think about which components you would drop and make a decision

## A1: I would drop PC6 since PC1-PC5 capture approx 91% of the variance based on the cumulative proportion shown in the summary.

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D2 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D2.1 <- cbind(D2, as.data.frame(D1$mean_correct))

# Q2: Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?
 
#A2:Not if we drop PC3 or PC5 since they have virtually no correlation with mean_correct

COR2 <- cor(D2.1)

#Correlogram 2: PCs and Mean Correct
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank", main = "Assistments PCs and Mean Correct")
```
## Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

# Q3:Now examine your components and try to come up with substantive descriptions of what some might represent?

# A3: PC1-PC6 Top 3 Variable Contributions:

#PC1 (mean_hint, mean_attempt, and problems_attempted) could represent the extent a student seeks/receives external feedback during a session 
fviz_contrib(pca, 
             choice = "var", 
             axes = 1, 
             top = 3,
             title = "PC1 Top 3 Variable Contributions")

# PC2 (prior_percent_correct, prior_prob_count) seems to represent a student's past performance in relation to how many problems the answered 
fviz_contrib(pca, 
             choice = "var", 
             axes = 2, 
             top = 3,
             title = "PC2 Top 3 Variable Contributions")

#PC3 (mean_confidence) mostly represents a student's confidence level 
fviz_contrib(pca, 
             choice = "var", 
             axes = 3, 
             top = 3,
             title = "PC3 Top 3 Variable Contributions")

#PC4 (prior_prob_count, mean_confidence) seems to represent confidence considering the number of problems a student answered already 
fviz_contrib(pca, 
             choice = "var", 
             axes = 4, 
             top = 3,
             title = "PC4 Top 3 Variable Contributions")

#PC5 (mean_attempt, problems_attempted) seems to reflect student persistence 
fviz_contrib(pca, 
             choice = "var", 
             axes = 5, 
             top = 3,
             title = "PC5 Top 3 Variable Contributions")

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

#Biplot 1: Assistments PCA
biplot(pca,
       xlabs = rep(".", nrow(D1.1)),
       cex = c(1,0.5),
       xlim = c(-0.3, 0.3),
       ylim = c(-0.3, 0.3),
       arrow.len = 0.05,
       expand = 2,
       main = "Assistments PCA")

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

#Caluclate composites
Composites <- as.matrix(D1.1) %*% pca$rotation
Composites <- cbind(Composites, as.data.frame(D1$mean_correct))
colnames(Composites) <- c("C1", "C2", "C3", "C4", "C5", "C6", "mean_correct")

#Correlogram 3: Composite Variables and Mean Correct 
COR2.1 <- cor(Composites)
corrplot(COR2.1, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank", main = "Composite Variables and Mean Correct")

```
# Part III  
## Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

```{r}

#Load Data
D3 <- read.csv("humor_data.csv")

##Question-based PCA:
#Select Questions
D3.1Q <- D3[1:32]

#Correlogram 4: Humor Questions (too difficult to read)
COR3.1Q <- cor(D3.1Q)
corrplot(COR3.1Q, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col=NULL, addCoefasPercent = TRUE, sig.level=0.50, insig = "blank", main = "Humor Questions")

#PCA of questions
pca3Q <- prcomp(D3.1Q, scale = TRUE)

#PCA Variance Plot 2: Humor Questions PCs
plot(pca3Q, type = "lines", main = "Humor Questions PCs")

#Examines eigenvalues and cumulative proportion of variance
pca3Q$rotation
summary(pca3Q)

#Bind humor style with PC1-PC6
D3.2 <- as.data.frame(pca3Q$x)
D3.2 <- D3.2[1:6]
D3.3 <- cbind(D3.2, as.data.frame(D3[33:36]))

#Correlogram 5: Humor Style and Question-based PCs
COR3.1Q <- cor(D3.3)
corrplot(COR3.1Q, order="original", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank", main = "Humor Style and Question-based PCs")

#Biplot 2: Humor Questions PCA
biplot(pca3Q,
       cex = c(1,0.5),
       arrow.len = 0.05,
       xlabs = rep(".", nrow(D3.3)),
       main = "Humor Questions PCA")

##Exploratory Analysis of Humor Style:
#Select humor style
D3.1HS <- D3[33:36]

#Correlogram 6: Humor Style (least correlation between affiliative and aggresive humor)
D3.1HS <- scale(D3.1HS, center = TRUE)
COR3.1HS <- cor(D3.1HS)
corrplot(COR3.1HS, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank", main = "Humor Style")

#PCA of humor style
pca3HS <- prcomp(D3.1HS, scale = TRUE)

#PCA Variance Plot 3: Humor Styles PCs
plot(pca3HS, type = "lines")

#Examines eigenvalues and cumulative proportion of variance
pca3HS$rotation
summary(pca3HS)

#Biplot 3: Humor Style PCA
biplot(pca3HS,
       cex = c(1,0.5),
       arrow.len = 0.05,
       xlim = c(-0.2, 0.15),
       xlabs = rep(".", nrow(D3.3)),
       main = "Humor Style PCA")
```

