---
title: "Principle Component Analysis"
output: html_document
---
#Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:
=======
title: "Principle Component Aanalysis"
output: html_document
---
#Data
The data you will be using comes from teh Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

#Start by uploading the data
```{r}
#setwd("C:/Users/Magdalena Bennett/Dropbox/PhD Columbia/Fall 2016/Core Methods in EDM/assignment4")

D1 <- read.csv("Assistments-confidence.csv")
names(D1)

#We remove the ID, because we are not going to use it.
D1 <- D1[,-1]
names(D1)
```

#Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)

library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```

- The diagonal of the correlation plot, shows that there is a 100\% correlation between a variable and itself (which is fairly obvious). The size of the circles (and the intensity of the color) will graphically show us the magnitude of the correlation, while the color of the circle will show us whether the correlation is positive (blueish) or negative (redish). In this case, we can observe that *mean_hint* has a strong negative correlation to *mean_correct*, and a slight negative correlation with *prior_percent_correct*. The variable *mean_hint* also seems to be slighlty positively correlated to *median_attempt*, *problems_attempted*, and *prior_prob_count*. Additionally, *mean_attempt* is positively correlated with *prior_percent_correct* and *problems_attempted*. Other variables do nos seem to be significantly correlated.

#Create a new data frame with the mean_correct variables removed

```{r}
D2 <- D1[,-which(names(D1)=="mean_correct")]

names(D2)

```

#Create a new data frame with the mean_correct variables removed

```{r}
#The, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

#Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
```

#Although the algorithm does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can lot this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")

#We can also get the cummulative variation that they explain:

var <- pca$sdev^2
cv <- var/sum(var)

#Percentage of variation each component explain
plot(cv, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')

#Cummulative percentage of variation each component explains
plot(cumsum(cv), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')

```

#Think about which components you would drop and make a decision

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)
  
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
  
```

- We can see that if we use the principal components, by constructions they won't be correlated (because they have to be orthogonal). In this case, we can se that *mean_correct* is positively correlated (with a moderate magnitude) with the first and second principal component. As well, it appears to have a moderate negative correlation with the sixth principal component. In that sense, we probably are not losing much information by dropping *mean_correct* as it is being somewhat captured by the 1st, 2nd, and 6th principal component. This can also be observed in our first correlation plot, where *mean_correct* was moderately correlated with *prior_percent_correct* and highly correlated with *mean_hint*, which means that those variables contain a relevant portion of the information in the *mean_correct* variable.

- To check this, we can add the variable *mean_correct* and run PCA again:

```{r}

pca2 <- prcomp(D1, scale = TRUE)

summary(pca2)

plot(pca2, type = "lines")

var <- pca2$sdev^2
cv <- var/sum(var)

#Percentage of variation each component explain
plot(cv, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')

#Cummulative percentage of variation each component explains
plot(cumsum(cv), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

- Adding one more component (even though it is not directly comparable to using 6 components) appears not to make much difference, as the seventh PC explains very little of the variance. This would support our idea that we would not lose much information by not including *mean_correct*

- I assumed theat mean_correct ws just another characteristic, but if it is the outcome that we are interested in, then the same logic applies. Dropping the last PC because it explains little variance would probably be a mistake, as it is strongly correlated with the outcome of interest.

#Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

#sweep(loadings, 2, colSums(aload), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

sweep(loadings, 2, colSums(loadings), "/")

#Now examine your components and try to come up with substantive descriptions of what some might represent?

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  
par(mar = c(5,4,3,3) + .1, las = 1, cex = 0.8)

biplot(pca)

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

```

- We can see that the first principal component weights *mean_hint*, *mean_attempt*, and *problems_attempted* more heavily.This could potentially be a component that captures the overall "effort" of the student during the test, including the outside help (hints).

- The second PC weights heavily on *prior_percent_correct*, so it's a component that captures the previous success of the student.

- The third PC is weights more on *mean_confidence*, so this component reflects how confident the student is on their answers.

- The fourth PC weights more the prior number of problems the student did and also their confidence. This component probably reflects the effort of the student prior to the test.

- The fifth PC weights *mean_attemtp* and *problems_attempted*, which is similar to the first one, but does not consider the number of hints in this case, so it's like a "thanks for trying" component.

- Finally, the sixth PC seems to sort of weight all the variables (except *mean_confidence*), and more heavily *mean_hint*. So it could be a measure of "outside help".

#Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

```{r}
#Load the data
humor <- read.csv("humor_data.csv")

#First, we will extract only the questions
d_humor <- humor[,-(33:39)]

#Now, we want to see if we can recreate the 4 dimensions they captured:

#We also scale it and center it:
pcah <- prcomp(d_humor, scale. = TRUE)

pcah

biplot(pcah)

```
- In the biplot, we can clearly see that the questions take clearly four directions, that could be associated to the "types of humor" that the researchers are talking about.
