---
title: "Assignment 4: K Means Clustering"
Author: Xingyi Xie
---

In this assignment we will be applying the K-means clustering algorithm we looked at in class. At the following link you can find a description of K-means:

https://www.cs.uic.edu/~wilkinson/Applets/cluster.html


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

Now, upload the file "Class_Motivation.csv" from the Assignment 4 Repository as a data frame called "K1""
```{r}


K1 <- read.csv("Class_Motivation.csv", header = TRUE)

K1b <- gather(K1, week, measure, 2:6)

plot(as.factor(K1b$week), K1b$measure)

```

This file contains the self-reported motivation scores for a class over five weeks. We are going to look for patterns in motivation over this time and sort people into clusters based on those patterns.

But before we do that, we will need to manipulate the data frame into a structure that can be analyzed by our clustering algorithm.

The algorithm will treat each row as a value belonging to a person, so we need to remove the id variable.

```{r}

K2 <- select(K1, 2:6)

```

It is important to think about the meaning of missing values when clustering. We could treat them as having meaning or we could remove those people who have them. Neither option is ideal. What problems do you foresee if we recode or remove these values? Write your answers below:

Answer: In my opinion, when we remove those people who have them, the sample size will decrease a lot because not every individual has all meaningful data. 
Secondly, if we recode those missing values, it will also affect our statistical assumption, because we don't know where these missing values come and for what reasons.  In this case, there are some solutions by R to deal with missing values, like multiple imputation by package "mice" (I saw this in <R in action>). 
Therefore, I agree with that neither solution of them is ideal. 


We will remove people with missing values for this assignment, but keep in mind the issues that you have identified.


```{r}

K3 <- na.omit(K2) 
K3 <- K2

K3[is.na(K3)] <- 0


```

Another pre-processing step used in K-means is to standardize the values so that they have the same range. We do this because we want to treat each week as equally important - if we do not standardise then the week with the largest range will have the greatest impact on which clusters are formed. We standardise the values by using the "scale()" command.

```{r}

K3 <- scale(K3) 

```


Now we will run the K-means clustering algorithm we talked about in class. 
1) The algorithm starts by randomly choosing some starting values 
2) Associates all observations near to those values with them
3) Calculates the mean of those clusters of values
4) Selects the observation closest to the mean of the cluster
5) Re-associates all observations closest to this observation
6) Continues this process until the clusters are no longer changing

Notice that in this case we have 5 variables and in class we only had 2. It is impossible to vizualise this process with 5 variables.

Also, we need to choose the number of clusters we think are in the data. We will start with 2.

```{r}

fit1a <- kmeans(K3, 2)
fit1b <- kmeans(K3, 2)
fit1c <- kmeans(K3, 2) 

fit1b$cluster

K4 <- data.frame(K3, fit1a$cluster, fit1b$cluster, fit1c$cluster)

fit1a$withinss
fit1b$withinss
fit1c$withinss

fit1a$tot.withinss
fit1b$tot.withinss
fit1c$tot.withinss

fit1a$betweenss
fit1b$betweenss
fit1c$betweenss

K4 <- data.frame(K3, fit1a$cluster)


names(K4) <- c("1", "2", "3", "4", "5", "cluster") 
```

Now we need to visualize the clusters we have created. To do so we want to play with the structure of our data. What would be most useful would be if we could visualize average motivation by cluster, by week. To do this we will need to convert our data from wide to long format. Remember your old friends tidyr and dplyr!

First lets use tidyr to convert from wide to long format.
```{r}

K5 <- gather(K4, "week", "motivation", 1:5)
```

Now lets use dplyr to average our motivation values by week and by cluster.

```{r message=FALSE, warning=FALSE}
K6 <- K5 %>% 
  group_by(week, cluster) %>% 
  summarise(avg = mean(motivation))


```

Now it's time to do some visualization: 

https://www.cs.uic.edu/~wilkinson/TheGrammarOfGraphics/GOG.html

And you can see the range of available graphics in ggplot here:

http://ggplot2.tidyverse.org/reference/index.html

We are going to create a line plot similar to the one created in this paper about school dropout [Bowers, 2010](http://pareonline.net/pdf/v15n7.pdf). It will have motivation on the Y-axis and weeks on the X-axis. To do this we will want our weeks variables to be treated as a number, but because it was created from a variable name it is currently being treated as a character variable. You can see this if you click on the arrow on the left of K6 in the Data pane. Week is designated by "chr". To convert it to numeric, we use the as.numeric command. 

Likewise, since "cluster" is not numeric but rather a categorical label we want to convert it from an "integer" format to a "factor" format so that ggplot does not treat it as a number. We can do this with the as.factor() command.

```{r}

K6$week <- as.numeric(K6$week)

K6$cluster <- as.factor(K6$cluster)

```

Now we can plot our line plot using the ggplot command, "ggplot()".

- The first argument in a ggplot is the dataframe we are using: K6
- Next is what is called an aesthetic (aes), the aesthetic tells ggplot which variables to use and how to use them. Here we are using the variables "week" and "avg" on the x and y axes and we are going color these variables using the "cluster" variable
- Then we are going to tell ggplot which type of plot we want to use by specifiying a "geom()", in this case a line plot: geom_line()
- Finally we are going to clean up our axes labels: xlab("Week") & ylab("Average Motivation")

```{r}

ggplot(K6, aes(week, avg, colour = cluster)) + geom_line() + xlab("Week") + ylab("Average Motivation")

```

What patterns do you see in the plot?

### Answer: People of Cluter 1 and Cluster 2 have different pattern through the week. Cluster2 people has stronger average motivation than Cluster1 people and the motivation of cluster 2 achieves highest in Week 1 and week 3. However, compared to cluster 2, cluster 1 people has lowest average motivation value in week 1 and 3.

It would be useful to determine how many people are in each cluster. We can do this easily with dplyr.

```{r}
K7 <- count(K4, cluster)
```

Look at the number of people in each cluster, now repeat this process for 3 rather than 2 clusters. Which cluster grouping do you think is more informative? Write your answer below:


```{r Part 1 cluster 3, message=FALSE, warning=FALSE}

K3 <- na.omit(K2) 
K3 <- K2

K3[is.na(K3)] <- 0


fit1a <- kmeans(K3, 3)
fit1b <- kmeans(K3, 3)
fit1c <- kmeans(K3, 3) 

fit1b$cluster

K4 <- data.frame(K3, fit1a$cluster, fit1b$cluster, fit1c$cluster)

fit1a$withinss
fit1b$withinss
fit1c$withinss

fit1a$tot.withinss
fit1b$tot.withinss
fit1c$tot.withinss

fit1a$betweenss
fit1b$betweenss
fit1c$betweenss

K4 <- data.frame(K3, fit1a$cluster)

names(K4) <- c("1", "2", "3", "4", "5", "cluster") 

K5 <- gather(K4, "week", "motivation", 1:5)

K6 <- K5 %>% 
  group_by(week, cluster) %>% 
  summarise(avg = mean(motivation))

K6$week <- as.numeric(K6$week)

K6$cluster <- as.factor(K6$cluster)
ggplot(K6, aes(week, avg, colour = cluster)) + geom_line() + xlab("Week") + ylab("Average Motivation")
K7 <- count(K4, cluster)
```

### Answer: In my opinion, clustering data into 2 clusters is more informative compared to 3 clusters. Because we can see the different pattern according to this graph obviously. 


## Part II

Using the data collected in the HUDK4050 entrance survey (HUDK4050-cluster.csv) use K-means to cluster the students first according location (lat/long) and then according to their answers to the questions, each student should belong to two clusters.

```{r}
p2 <- read.csv("HUDK405020-cluster.csv")
L1 <- p2[,c(2,3)]
set.seed(121)
fit1a <- kmeans(L1, 2)
fit1b <- kmeans(L1, 2)

fit1a$cluster

L2 <- data.frame(L1, fit1a$cluster)


names(L2) <- c("Lat", "Long", "clusterLoc") 

ggplot(L2, aes(Lat,Long, fill = factor(clusterLoc))) + geom_boxplot() + xlab("Lat") + ylab("Lon")+theme_bw()
```


According to answers to questions
```{r}
k3 <- p2[,c(4:9)]
set.seed(121)
fit1a <- kmeans(k3, 2)
fit1b <- kmeans(k3, 2)

fit1a$cluster

K4 <- data.frame(k3, fit1a$cluster)

names(K4)[[7]] <- "cluster"

K5 <- gather(K4, "Questions","Answers", 1:6)

K6 <- K5 %>% group_by(Questions, cluster) %>% summarise(avg = mean(Answers))

K6$cluster <- as.factor(K6$cluster)
ggplot(K6, aes(Questions, avg, colour = cluster)) + geom_point() + xlab("Questions") + ylab("Average score")
K7 <- count(K4, cluster)
```


##Part III

Create a visualization that shows the overlap between the two clusters each student belongs to in Part II. IE - Are there geographical patterns that correspond to the answers? 


```{r}
D1 <- cbind(K4,L2)
pairs(D1)

chisq.test(D1$cluster,D1$clusterLoc)
tableone::CreateCatTable("cluster","clusterLoc",data = D1)

library(vcd)
P1 <- structable(D1$cluster ~ D1$clusterLoc)
mosaic(P1, shade=TRUE, legend=TRUE) 
```

### Answer: We didn't observe any geographical patterns that correspond to the answers. For Chi-square test (independence test), there is no significant relationships between geographical pattern and students' answers. Moreover, we can testify this conclusion from picture as belows.


## Please render your code as an .html file using knitr and Pull Resquest both your .Rmd file and .html files to the Assignment 3 repository.

