
 title: "Assignment 4: K Means Clustering"
 title: 'Assignment 4: K Means Clustering'
 ---

 In this assignment we will be applying the K-means clustering algorithm we looked at in class. At the following link you can find a description of K-means:

 https://www.cs.uic.edu/~wilkinson/Applets/cluster.html

 ## Part I

 ```{r}
 library()
 library(tidyr)
 library(dplyr)
 library(ggplot2)
 ```

 Now, upload the file "Class_Motivation.csv" from the Assignment 4 Repository as a data frame called "K1""
 ```{r}

 K1 <- read.csv(...)
 K1 <- read.csv("Class_Motivation.csv")

 ```

 @@ -26,7 +29,7 @@ The algorithm will treat each row as a value belonging to a person, so we need t

 ```{r}

 K2 <- 
 K2 <- K1[,-1]

 ```

 @@ -47,7 +50,7 @@ Another pre-processing step used in K-means is to standardize the values so that

 ```{r}

 K3 <- 
 K3 <- scale(K3)

 ```

 @@ -66,7 +69,7 @@ Also, we need to choose the number of clusters we think are in the data. We will

 ```{r}

 fit <- 
 fit <- kmeans(K3,centers = 2)

 #We have created an object called "fit" that contains all the details of our clustering including which observations belong to each cluster.

 @@ -76,7 +79,7 @@ fit <-

 #We can also attach these clusters to the original dataframe by using the "data.frame" command to create a new data frame called K4.

 K4
 K4 <- data.frame(K3, cluster = fit$cluster)

 #Have a look at the K4 dataframe. Lets change the names of the variables to make it more convenient with the names() command.

 @@ -95,7 +98,7 @@ Now lets use dplyr to average our motivation values by week and by cluster.

 ```{r}

 K6 <- K5 %>% group_by(week, cluster) %>% summarise(K6, avg = mean(motivation))
 K6 <- K5 %>% group_by(week, cluster) %>% summarise( avg = mean(motivation))

 ```

 @@ -113,9 +116,9 @@ Likewise, since "cluster" is not numeric but rather a categorical label we want

 ```{r}

 K6$week <- 
 K6$week <- as.factor(K6$week)

 K6$cluster <- 
 K6$cluster <- as.factor(K6$cluster)

 ```

 @@ -128,13 +131,13 @@ Now we can plot our line plot using the ggplot command, "ggplot()".

 ```{r}

 ggplot(K6, aes(week, avg, colour = cluster)) + geom_line() + xlab("Week") + ylab("Average Motivation")
 ggplot(K6, aes(week, avg, colour = cluster, group = cluster)) + geom_line() + xlab("Week") + ylab("Average Motivation")

 ```

 What patterns do you see in the plot?


 *Two wave line*

 It would be useful to determine how many people are in each cluster. We can do this easily with dplyr.

 @@ -144,18 +147,65 @@ K7 <- count(K4, cluster)

 Look at the number of people in each cluster, now repeat this process for 3 rather than 2 clusters. Which cluster grouping do you think is more informative? Write your answer below:

 ##Part II
 ```{r}
 fit_3<-kmeans(K3, 3)
 ```


 ```{r}
 K8 <- data.frame(K3, cluster  = fit_3$cluster)
 ```

 ```{r}
 K9 <- gather(K8,"week","motivation",1:5)
 ```


 ```{r}
 K10 <- K9 %>% group_by(week,cluster) %>% summarise(avg=mean(motivation))
 ```


 ```{r}
 K10$week<-as.factor(K10$week)
 K10$cluster<-as.factor(K10$cluster)

 ggplot(K10,aes(week,avg,colour=cluster,group = cluster))+geom_line()+xlab("week")+ylab("Average Motivation")
 ```
 *I think 2 clusters are better because they have less interactions*

 ## Part II

 Using the data collected in the HUDK4050 entrance survey (HUDK4050-cluster.csv) use K-means to cluster the students first according location (lat/long) and then according to their answers to the questions, each student should belong to two clusters.
 ```{r}
 D1 <- read.csv("HUDK405020-cluster.csv")
 D2 <- D1[,c(2,3)]
 D3 <- scale(D2)
 km1 = kmeans(D3,centers = 3)
 D4 <- D1[,c(4:9)]
 D5 <- scale(D4)
 km2 = kmeans(D5,centers=4)
 D6 <- data.frame(D2,D4,cluster1=as.vector(km1$cluster),cluster2=as.vector(km2$cluster))
 ```

 ##Part III
 ## Part III

 Create a visualization that shows the overlap between the two clusters each student belongs to in Part II. IE - Are there geographical patterns that correspond to the answers? 

 ```{r}

 library(sp)
 library(maps)
 library(maptools)
 ```

 ```{r}
 mp=NULL

 mapworld=borders('world',colour = 'grey100',fill='white')

 ggplot()+mapworld+ylim(-60,90)+geom_point(D6,mapping=aes(x=long,y=lat,shape=as.factor(cluster1),col=as.factor(cluster2)))
 ```
 *There is no relation between their locations and their answers*

 ## Please render your code as an .html file using knitr and Pull Resquest both your .Rmd file and .html files to the Assignment 3 repository.
