---
title: "DA_Assignment4_Final"
author: "DA"
date: "November 15, 2018"
output: html_document
---
```{r}
#Part 1
#create new data frame
D1 <- read.csv("Assistments-confidence.csv", header = TRUE)
#remove ID varibale
library(dplyr)
D1 <- select(D1, prior_prob_count:mean_confidence)
library(corrplot)
#Generate pairwise coorelation
COR <- cor(D1)
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
#create a new data frame by removing mean_correct variable
D2 <- D1[,-4]
#data scaled for easier interpretation
D2 <- scale(D2, center = TRUE)
#Run PCA in new data frame
pca <- prcomp(D2, scale = TRUE)
#print standard deviations
pca$sdev
#change to variance by squaring it
pca$sdev^2
#following summary will give the proprtion of variance accounted for by each component
summary(pca)
plot(pca, type = "lines")
#Glancing at the proportion of variance, PC5 and PC6 should be dropped. Note that PC5 accounts for 12 percent variance while PC6 accoutns for only 8 percent. 
```

```{r}
#Part II
D3 <- as.data.frame(pca$x)
#attach the mean_correct abck to the new data frame by useing cbind function
D4 <- cbind(D3, as.data.frame(D1$mean_correct))
#re-run the scatterplots
COR2 <- cor(D4)
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
pca$rotation
#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component
loadings <- abs(pca$rotation)
#sweep() computes each row as a proportion of the column
sweep(loadings, 2, colSums(loadings)
#with Dplyr
library(tidyr)
loadings1 <- tbl_df(loadings)
Variables <- c("prior_prob_count", "prior_percent_correct", "problems_attempted", "mean_hint", "mean_attempt", "mean_confidence ") 
loadings2 <- cbind(Variables, loadings1)
loadings3 <- loadings2 %>% mutate(PC1 = PC1/sum(PC1)) %>% mutate(PC2 = PC2/sum(PC2)) %>% mutate(PC3 = PC3/sum(PC3)) %>% mutate(PC4 = PC4/sum(PC4)) %>% mutate(PC5 = PC5/sum(PC5)) %>%  mutate(PC6 = PC6/sum(PC6)) %>% print
#compute biplot
biplot(pca)
#PC1 contributes the most to mean_hint, PC2 to prior_percent correct, PC3 to mean_confidence, PC4 to prior_percent_correct, Pc_5 to mean attempt and PC6 to mena_hint. 
D3<- as.matrix(D2) %*% pca$rotation
D4 <- cbind(D3, as.data.frame(D1$mean_correct))
COR3 <- cor(D4)
corrplot(COR3, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

```

```{r}
#part III
A1 <- read.csv("humor_data.csv", header=TRUE)
(A1 <- A1[,1:32])
A2 <- scale(A1, center = TRUE)
pca2 <- prcomp(A2, scale = TRUE)
#standard deviation and variance
pca2$sdev
pca2$sdev^2
summary(pca2)
plot(pca2, type = "lines")
biplot(pca2)
loadings4 <- abs(pca2$rotation)
loadings5 <- sweep(loadings4, 2, colSums(loadings4), "/")
A3 <- as.matrix(A2) %*% pca2$rotation
COR4 <- cor(A3)
corrplot(COR4, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")


